The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-11-10 paper/simplest_effects.tex
paper WIP commit

@@

Since the effects of binarity are most pronounced when the two components are 
similar, we begin by considering a universe in which all planets are 
identical, and all stars are either single or twin stars with otherwise 
identical physical properties.
The occurrence rate density at a planet radius $r$, stellar radius $R$, and 
semimajor axis $a$ for this model is
\begin{equation}
\Gamma(r,R,a) = \sum_i w_i \Lambda_i \delta^3(r_p,R_\star,a_p),
\label{eq:model1_occ_rate_density}
\end{equation}
for $r_p$ and $R_\star$ some fixed planet and stellar radii, $a_p$ a fixed 
semi-major axis, and $\delta$ the Dirac delta function, whose compact
form will be used for brevity.
The occurrence rate over any interval that includes $r_p$, $R_\star$, and 
$a_p$ is
\begin{equation}
\Lambda = \sum_i w_i \Lambda_i = \frac{ \sum_i N_i \Lambda_i }{N_{\rm tot}}.
\end{equation}
The rate is zero over intervals that do not.

@@

\paragraph{What does an observer ignoring binarity infer?} 
An observer who ignores binarity assumes a detection efficiency 
$\tilde{Q}=Q_0$,
measures a detected planet rate density $\tilde{\Gamma}$, 
and infers an apparent rate density $\Gamma_a$.
Analogous to Eq.~\ref{eq:detected_rate_density},
\begin{equation}
\tilde{\Gamma} = \Gamma_a \tilde{Q}.
\end{equation}
Accounting for dilution, and assuming $q=1$, the apparent rate density is
\begin{equation}
\Gamma_a = 
w_a \Lambda_0 \delta^3(r_p, R_\star, a_p) +
w_b (\Lambda_1 Q_{c,1} + \Lambda_2 Q_{c,2}) 
				\delta^3(r_p/\sqrt{2}, R_\star, a_p),
\end{equation}
for $w_a = N_0/(N_0+N_1)$, and $w_b = N_1/(N_0+N_1)$.
This observer miscounts the number of total searched stars, does not correct 
for incompleteness, and misclassifies the planetary radii because of dilution.

>>>>

\section{Idealized Models of Transit Surveys}

@@

Since the effects of binarity are most pronounced when the two components are 
similar, we begin by considering a universe in which all planets are 
identical, and all stars are identical except that some fraction of them exist 
in binaries.

Expressed mathematically, from Eqs.~\ref{eq:rate_density_marginalized}
and~\ref{eq:rate_density_shape} the occurrence rate density at a planet radius 
$r$ can be written
\begin{equation}
\Gamma(r) = \delta(r_p) \times
\frac{
    N_0 Z_0 +
    N_1 Z_1 +
    N_2 Z_2
}{N_{\rm tot}},
\label{eq:model_1_rate_density}
\end{equation}
where $\delta(r_p)$ is the Dirac delta function, zero except at the true 
planet radius, $r_p$.
The occurrence rate over any interval that includes $r_p$ is then
\begin{equation}
\Lambda|_{r_p} = \frac{
    N_0 Z_0 +
    N_1 Z_1 +
    N_2 Z_2 
}{N_{\rm tot}},
\label{eq:model_1_rate}
\end{equation}
and the rate is zero over intervals that do not include $r_p$.

We return to our group of binarity-ignoring astronomers. They do 
not know the true planet population~--~they would like to discover it!
In their signal-to-noise limited transit survey, they select stars 
that they think can yield transit detections.
Since the noise is Poissonian, they assume
\begin{equation}
\frac{{\rm signal}}{{\rm noise}}
\propto \frac{(r/R)^2}{F^{-1/2}}
\propto F^{1/2}
\propto L_{\rm sys}^{1/2} d^{-1}, \quad ({\rm incorrectly\ assumed})
\end{equation}
for $R$ the (constant) stellar radius, $F$ the photon flux, $L_{\rm sys}$ the 
luminosity of a system, and $d$ its distance from us.
At fixed planet radius, semimajor axis, stellar radius, and stellar luminosity,
a constant signal-to-noise floor yields a maximum detectable 
distance (Pepper et al 2003, Pepper and Gaudi 2005).
The maximum distance out to which our binarity-ignoring astronomers 
select stars, $d_{\rm sel}$, thus scales as $L_{\rm sys}^{1/2}$.

The single stars have luminosity $L_1$, and the twin binaries have luminosity 
$2L_1$.
Thus the twin binaries are selected out to a distance $\sqrt{2}$ times 
that of single stars.
This is a bad move; in reality the transit signal for any planet in a 
twin binary will be diluted by a factor of two:
\begin{equation}
\frac{{\rm signal}}{{\rm noise}}
\propto \frac{\mathcal{D} (r/R)^2}{F^{-1/2}}
\propto \mathcal{D} F^{1/2}
\propto \mathcal{D} L_{\rm sys}^{1/2} d^{-1}, \quad ({\rm true})
\end{equation}
where the dilution is $\mathcal{D} \equiv L_{\rm host}/L_{\rm sys}$, for 
$L_{\rm host}$ the the planet host's luminosity.
This means that the actual maximum searchable distance for binaries is 
$1/\sqrt{2}\times$ that of single stars.
The situation is illustrated in Fig.~\ref{fig:model_1_volumes}: only one in 
eight selected stars in binaries are truly searchable.

\paragraph{What do the observers ignoring binarity infer?} 
The binarity-ignoring observers assume that every point on the sky with flux 
above some minimum are searchable.
They measure a detected planet rate density, and infer an apparent rate 
density by correcting for $p_{\rm tra}$.
The apparent rate density is
\begin{equation}
\Gamma_a(r) = 
\delta(r_p) Z_0 \frac{N_0}{N_0+N_1}  +
\delta\left(\frac{r_p}{\sqrt{2}}\right) 
(Z_1 p_{{\rm det},1} + Z_2 p_{{\rm det},2}) \frac{N_1}{N_0+N_1},
\label{eq:model_1_apparent_rate_density}
\end{equation}
where $p_{{\rm det},1}$ ($p_{{\rm det},2}$) is the probability that a selected 
primary (secondary) is searchable.
For this example, $p_{{\rm det},1}=p_{{\rm det},2}=1/8$.

The true rate density (Eq.~\ref{eq:model_1_rate_density}) and the
apparent rate density (Eq.~\ref{eq:model_1_apparent_rate_density})
differ in that
\begin{enumerate}
\item The total number of selected stars, $N_{\rm tot} = N_0+N_1+N_2$, was 
miscounted.
%
\item The detection efficiency was incorrectly assumed to be $1$ for all 
selected stars. In reality, only one in eight binaries were searchable.
%
\item The inferred radii in binary systems are all $\sqrt{2}$ too small.
\end{enumerate}

We show the resulting occurrence rates over bins in planet radius in 
Fig.~\ref{fig:errcases_model_1}.

%NOTE: YOU CAN BUTCHER THE FUCK OUT OF THIS SECTION. JNW WANTS EACH CASE TO BE 
%EXPLAINED, INDIVIDUALLY. DROP ALL THE "i" SUBSCRIPTS B/C THEY ARE APPARENTLY 
%CONFUSING.

@@
